---
title: "Simulations for ICES categories 4-6"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repository ([GA_MSE_cat456](https://github.com/shfischer/GA_MSE_cat456)) is a mirror of [GA_MSE_HR](https://github.com/shfischer/GA_MSE_HR), [GA_MSE](https://github.com/shfischer/GA_MSE), and [GA_MSE_PA](https://github.com/shfischer/GA_MSE_PA) with the `cat456` branch displayed as default branch.

## Introduction

This repository contains the code for an exploration of the ICES approach for categories 4-6. It includes the typical ICES harvest control rule (constant catch with PA buffer) and a version with a conditional PA buffer based on length data (including optimisation with a genetic algorithm). The simulation is based on the Fisheries Library in R ([FLR](http://www.flr-project.org/)) and the Assessment for All (a4a) standard MSE framework ([`FLR/mse`](github.com/FLR/mse)) developed during the Workshop on development of MSE algorithms with R/FLR/a4a ([Jardim et al., 2017](https://ec.europa.eu/jrc/en/publication/assessment-all-initiativea4a-workshop-development-mse-algorithms-rflra4a)).

The simulation framework is based on the ([GA_MSE_HR](https://github.com/shfischer/GA_MSE_HR)) repository which explores the use of harvest rates and contains the code for the publication:

> Fischer, S. H., De Oliveira, J. A. A., Mumford, J. D., and Kell, L. T. (2022). Exploring a relative harvest rate strategy for moderately data-limited fisheries management. ICES Journal of Marine Science. 79: 1730-1741. https://doi.org/10.1093/icesjms/fsac103.

The operating models provided as an input are those from the repository [shfischer/wklifeVII](https://github.com/shfischer/wklifeVII) as described in:

> Fischer, S. H., De Oliveira, J. A. A., and Laurence T. Kell (2020). Linking the performance of a data-limited empirical catch rule to life-history traits. ICES Journal of Marine Science, 77: 1914-1926. https://doi.org/10.1093/icesjms/fsaa054.


## Repository structure

The root folder contains the following R scripts:

* `OM.R`: This script creates the operating models (OMs),
* `funs_OM.R` contains functions and methods used for the creation of the operating models 
* `funs.R` contains functions for running the MSE such as the OM and MP modules,
* `funs_GA.R` contains the function used in the optimisation procedure,
* `run_ms_cat456.R` is an R script for running MSE projections and is called from a job submission script
* `run*.pbs` are job submission scripts which are used on a high performance computing cluster and call `run_ms.R`
* `analysis_cat456.R` is for analysing the results

The following input files are provided:

* `input/stocks.csv` contains the stock definitions and life-history parameters
* `input/brps.rds` contains the FLBRP objects which are the basis for the OMs

Summarised outputs are provided in:

* `output/`




## R, R packages and version info

This is branch `cat456` in which R and R packages have been updated to R 4.2 (4.3 should also work):
```{r, eval = FALSE}
> sessionInfo()
R version 4.2.3 (2023-03-15 ucrt)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 19043)
```

The package versions and their dependencies are recorded with the R package [renv](https://rstudio.github.io/renv/) and stored in the file [renv.lock](https://github.com/shfischer/GA_MSE_HR/blob/harvest_rate_R4.2/renv.lock). The exact package version can be restored by cloning this repository, checking out the current branch (`cat456`), navigating into this folder in R (or setting up a project), installing the renv package

```{r, eval = FALSE}
install.packages("renv")
```

and calling

```{r, eval = FALSE}
renv::restore()
```

See [renv](https://rstudio.github.io/renv/) and the package documentation for details.

The framework is based on the Fisheries Library in R (FLR) framework and uses the [FLR packages](https://flr-project.org/) `FLCore`, `FLash`, `FLBRP`, `ggplotFL`, `mse`. See [renv.lock](https://github.com/shfischer/GA_MSE_HR/blob/cat456/renv.lock) for version details and sources.

The FLR package versions can also be installed manually with `remotes` (requires suitable tools to compile R packages):
```{r, eval = FALSE}
remotes::install_github(repo = "flr/FLCore", ref = "031f7742bca4b128f80ee3cc386c21d9c00ddcf7")
remotes::install_github(repo = "flr/FLash", ref = "1008c040b760f68df5101f8aafba28d9e61bba5e", INSTALL_opts = "--no-multiarch")
# INSTALL_opts = "--no-multiarch" to avoid issues in Windows
remotes::install_github(repo = "flr/FLBRP", ref = "5fa2b741d81f3d5c3e06b40cc221c98418d3a3c7", INSTALL_opts = "--no-multiarch")
remotes::install_github(repo = "shfischer/mse", ref = "18ce8c8d51362d5ced27b6c5978912f6da0ab5dc", INSTALL_opts = "--no-multiarch")
# and the GA package for running the genetic algorithm
remotes::install_github(repo = "shfischer/GA", ref = "48c1092437629b86a5310fa2873621621ff0b0e0")
```

For using MPI parallelisation, an MPI backend such as OpenMPI and the R packages `Rmpi` and `doMPI` are required.
